Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities

### 2021-03-19 docs/manuscript/main.tex
#4 cleans main.tex

\section*{Introduction}
The complexity of human movement arises from the process of mastering 
redundant (bio)mechanical degrees of freedom (DOF) to successfully 
accomplish any given motor task (Bernstein, 1967).
Such DOF are independent coordinates to uniquely 
describe the configurations of the system which provides flexibility 
and stability to adapt to a change environment conditions but 
that leads many possible combinations (Newell and Vaillancourt, 2001).
Hence, human movement complexity can be seen as a balance 
between the required DOF "to generate a stable (persistent) 
and flexible (variable) behavioural output in response to 
changing intentions and dynamic environmental
conditions" \cite{davids2003}. 
Consequently, one can see much variability in human movement even in the 
simplest of movements.  
While the analysis of movement of variability is becoming increasingly 
popular as a diagnostic tool or skill performance evaluation, 
there remain challenges in terms of defining 
the most appropriate methods and parameters to apply. 
In part, 
these challenges stem from the fact that the identification movement 
variability requires analysis of signals which are time-series of 
of $1-$dimension in $\mathbb{R}$ which are 
noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
Further problems arise from assigning a plausible locus of control to 
the movements, e.g., in order to determine whether variability is the 
result of deliberative control by the performer or whether it arises 
from exogenous or endogenous disturbances.  For this paper, our focus 
is on the analysis of signals; specifically, in terms of objectively 
quantifying variability of lower dimension signals using time-series analysis.

Methods for nonlinear time series analysis generally involve the estimation of the 
embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to 
reconstruct the state space,
where an $n$-dimensional is reconstructed state space using $1-$dimensional 
time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, 
frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
Key to the selection of these parameters is the need to preserve 
topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
An common approach to the construction of these state spaces 
involves Recurrence Plots (RPs), a graphical representation of black and white dots, 
which shows recurrent patterns of a $n$-dimensional system.
While RPs provide a human
interpretable picture of the system, these require further
analysis to allow the properties of that system to be quantified,
and so Recurrence Quantification Analysis (RQA) can be applied.
However, the estimation of the embedding parameters for
RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015})
There is no agreed method to estimate embedding parameters
for RQA and other nonlinear methods \cite{bradley2015} 
because time series are system-dependent, i.e., these rely
on the initial conditions, and on the configuration and behaviour of the system.  
This means that, unless one holds all of the influencing 
variables constant, embedding parameters computed for one 
instance may not apply to another instance.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Methods}
\subsection*{State Space Reconstruction}
The method of state space reconstruction \cite{packard1980} 
\cite{takens1981} has been applied in many disciplines 
\cite{aguirre2009, stergiou2011, frank2010, 
sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
based on uniform time-delay embedding methodology which is a simple 
matrix implementation that can reconstruct an unknown $d-$dimensional 
manifold $M$ from a scalar time series (e.g. one-dimensional 
time series in $\mathbb{R}$). A manifold, in this context, is a 
multidimensional curved surface within a space (e.g. a saddle) 
\cite{guastello-gregson2011}.

The use of a scalar time series is the main advantage of the method of state 
space reconstruction which in essence preserve dynamic invariants such as 
correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai 
entropy and detrended fluctuation analysis \cite{bradley2015, 
Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
However, selecting appropriate embedding parameters which are sued to apply the 
state space reconstruction is still an open challenge for which we present 
introductions for the methodologies to compute such embedding parameters 
With that in mind, in the following subsections, we describe in more detail 
the state space reconstruction theorem (RSSs), uniform time-delay embedding 
theorem (UTDE), false nearest neighbours (FNN) and average mutual 
information (AMI). 
%and other methodologies for state space reconstruction.

@@

Following the notation employed in \cite{casdagli1991, garland2016, gibson1992,
uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:

@@

where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and 
$M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an 
unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution 
function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th 
iteration of $f$ that corresponds to an initial position 
$s(0) \in M $ \cite{takens1981}.
%$f$ is a smooth dynamical system, where smooth means a 
%continuously differentiable system
%(e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, 
can be obtained with

@@

Reconstructed state space can then be described as an $n-$dimensional state 
space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where 
$\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ 
is the uniform time-delay embedding with a dimension embedding $m$
and delay embedding $\tau$ and
$ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation
of dimensionality (e.g. Principal Component Analysis, 

@@

With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, 
defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that 
$\boldsymbol{X}(t) = \Phi(s(t))$,
where $\Phi$ is a diffeomorphic map \cite{takens1981}
whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting
dimension of $M$ \cite{garland2016}.
% "Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$
% is called diffeomorphic if it is one-to-one correspondence and its inverse
% $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed 
space then a composition of functions represented with $F^t$ is induced on the
reconstructed state space determined:

@@

With this in mind, an embedding is defined as "a smooth one-to-one 
coordinate transformation with a smooth inverse" and the total reconstruction 
map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.

@@

		Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method (B), and
		Recurrence Plots and its patterns (C).

@@

For example, we compute the AMI for two time series:
A) the $x$ solution of the Lorenz system, and
B) a noise time series using a normal distribution with mean zero and 
standard deviation equal to one. From Fig~\ref{fig01}(b4, b8), it can then be
concluded that the amount of knowledge for any noise time series is zero 
for which the first minimum embedding parameter is $\tau_0=1$. 
On the contrary, the first minimum of the AMI for the chaotic time series 
is $\tau_0=17$ which is the value that maximize the independence 
between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
\cite{bradley2015}.
%%---------------------------------(FIGURE)-------------------------------------
Similarly as Cao's algorithm negatives, AMI's algorithm is not an
exception for negatives, which are worthwhile to mention for further 
investigations.
For instance, (i) is not clear why the choose of the first minimum of the AMI 
is the minimum delay embedding parameter \cite{kantz2003} and 
(ii) the probability distribution of the AMI function is computed
with the use of histograms which depends on a heuristic choice of number of bins
for which AMI depends on partitioning \cite{garcia2005e71}.

%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/methods/caoami/versions/drawing-v00.pdf}
%    \caption{
%	{\bf Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method.} 
%	(A) chaotic and (B) random time series.
%	(C, D) $E_1 (m)$ values and (E, F) $E_2(m)$ values 
%	with variations of $\tau$ values from one to twenty
%	for chaotic and random time series.
%	(G, H) AMI values where its first minimum value in the curve
%	is the minimum time delay embedding ($\tau_0$).
%	Code and data to reproduce the figure is available in \cite{srep2021}.
%        }
%    \label{fig:caoami}
%\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/methods/rps/versions/drawing-v00.pdf}
%    \caption{
%	{\bf Recurrence Plots and its patterns.} 
%	(A) State space of the Lorenz system with controlling parameters 
%	($\rho=28, \sigma=10, \beta=8/3$) where the point, $j$, in trajectory $X()$ which falls into the neighborhood 
%	(black circle) of a given point at $i$ is a recurrent point and is 
%	represented as a black dot in the recurrence plot at location 
%	$(i, j)$ or white otherwise.
%	(B) Recurrence plot using the 
%	three components of the Lorenz system and the RP with no embeddings 
%	and threshold $\epsilon=5$.
%	Time-series with its respective recurrence plot for:
%	(C) uniformly distributed noise,
%	(D) super-positionet harmonic oscillation 
%	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
%	(E) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
%	with a linearly increase term ($0.01*i$), and
%	(F) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
%	This figure is adapted from \cite{marwan2015}.
%	Code and data to reproduce the figure is available in \cite{srep2021}.
%	}
%    \label{fig:rps}
%\end{figure}
%%%---------------------------------(FIGURE)-------------------------------------

%*******************************************************************************
\subsection*{Recurrence Quantifications Analysis (RQA)}
Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate 
the density of recurrence points in RPs, then histograms of lengths for 
diagonal lines in RPs were studied by \cite{trulla1996} which were the 
introduction to the term recurrence quantification analysis (RQA) 
\cite{marwan2008}. RQA has been applied in many fields such as life science, 
engineering, physics, and others \cite{marwan2008}. Particularly in human 
movement to investigate noise and complexity of postural control 
\cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal 
coordination \cite{duran2017}. The success of RQA is not only due to its 
simple algorithmic implementation but also to its capacity to detect tiny 
modulations in frequency or phase which are not detectable using standard 
methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that 
RQA's metrics are quantitatively and qualitatively independe]nt of embedding 
dimension which is verified experimentally by \cite{iwanski1998}.
RQA metrics comprehend percentage of recurrence, percentage of determinism, 
ratio, Shannon entropy of the frequency distributions of the line lengths,
maximal line length and divergence, trend and laminariy 
\cite{marwan2007, marwan2015}. For this work, we considered only four 
RQA metrics, due to its consistency with our preliminary experiments, 
which are described below. Such metrics are computed the nonlinearTseries 
R package \cite{nonlinearTseries2016}.

@@

With that in mind, this work explores the sensitivity and robustness 
of the window size of time series, embedding parameters for RSS with UTDE 
and recurrence threshold for RP and RQA in order to gain a better insight 
into the underlying time series collected from inertial sensors in the 
context of human-humanoid imitation activities.
%Such changes of recurrence threshold values 
%can modify the patterns in RPs and therefore the values of RQA metrics.
%We therefore computed 3D surfaces to explore the sensibility and robustness of 
%embedding parameters and recurrence threshold in RQA  metrics. Following the 
%same methodology of computing 3D surfaces, we also considered variation of 
%window length size to present RQA metrics dependencies with embedding 
%parameters, recurrence thresholds and window length size.

%Choosing an appropriate recurrence threshold is crucial so as to get 
%meaningful representations in RP and RQA, however,

%Iwansky et al. \cite{iwanski1998} stated that patterns in Recurrence Plots and 
%metrics for RQA are independent of embedding dimension parameters, 
%however, that is not the case when using 
%different recurrence thresholds.

%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of 
%determining embedding parameters, their method consisted on creating 3D 
%surfaces with RQA metrics with an increase of embedding parameters 
%($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored 
%fluctuations and gradual changes in the 3D surfaces that provide information 
%about the embeddings. Much recently, Marwan et al. \cite{marwan2015} 
%created 3D surfaces for visual selection of not only embedding parameters 
%but also recurrence thresholds. Following same methodologies, we explored 
%the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR)
%using 3D surfaces by an unitary increase of the pair embedding 
%parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase 
%of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) 
%(Fig.~\ref{fig:topo_rqas}).

@@

\subsection*{Ethics}
All participants provided informed consent forms prior to participation 
in the experiment. Hence, the experiments were conducted in November 2016 and 
participants confirmed reading and understanding the participant information sheet of the 
experiments and were able to withdraw from the experiment at any time without giving
any reason. The design of the experiments adhered to the University of Birmingham
regulations, data were anonymised and videos were stored only on a 
personal computer in accordance with the Data Protection Act 1998. 
%Refer to Appendix D for further information about the ethics, 
%online participation information sheets and experiment check list.

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
two of which were attached to the right hand of the participant and the 
other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal
and vertical arm movements performed by the humanoid robot in the following 
conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) 
	and faster (HF) speed (Figure~\ref{fig02}a1), and
(ii) ten repetitions of vertical arm movement at normal (VN) 
	and faster (VF) speed (Figure~\ref{fig02}a3).
The normal and faster speed of arm movements is defined by the duration 
in number of samples of one repetition of NAO's arm movements. 
We select NAO's arm movements duration to distinguish between normal and 
faster arm movements as NAO's movements have less variation 
for such repetitive movements. 
The duration for one repetition of the horizontal 
arm movement at normal speed, HN, is 5 seconds considering 
that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed 
in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
which is around 300 samples of data. For vertical arm movement at 
faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal 
and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series
for axes Z and Y of the gyroscope sensors with four window lengths: 
2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
and 15-sec (750-samples).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
%    \caption{
%	{\bf Time series duration of horizontal and vertical arm movements.}
%    		Time series of smoothed data from gyroscope sensor
%		for different speed arm movements performed by NAO:
%		(A) Horizontal Normal arm movement (HN),
%		(B) Horizontal Faster arm movement (HF),
%		(C) Vertical Normal arm movement (VN) and
%		(D) Vertical Faster arm movement (VF).
%		Additionally,
%		(A) shows window sizes for 2-seconds (100 samples),
%		5-seconds (250 samples), 10-seconds (500 samples)
%		and 15-seconds (750 samples)
%		which are also presented in (B), (C) and (D).
%		Code and data to reproduce the figure is available in \cite{srep2021}.
%        }
%	\label{fig:sts}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------

\subsection*{Data collection from inertial measurement units} 
\label{sec:experiment:subsec:imu}
To give insight to the research questions, 
we considered various conditions of time series collected for this work: 
\begin{itemize}
\item Three levels of smoothness for normalised data 
	(e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg 
	stands for Savitzky-Golay filter with two different filter lengths (29 and 159) 
	and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal}
	and zmuv is zero mean unit variance.
\item four arm movement activities with two velocities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}
%After the data collection, raw time series were windowed, normalised and smoothed. 
Due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}C).
%%%---------------------------------(FIGURE)-------------------------------------
%\begin{figure}[ht]
%\centering
%\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
%    	\caption{
%	{\bf Time series for horizontal and vertical arm movements.}
%		(A/D) raw-normalised (sg0zmuv),
%		(B/E) normalised-smoothed 1 (sg1zmuv) and
%		(C/F) normalised-smoothed 2 (sg2zmuv).
%		Time series are only for three participants
%		($p01$, $p02$, and $p03$)
%		for horizontal and vertical arm movements in normal
%		and faster velocity (HN, HF, VN, VF)
%		with the normalised GyroZ or GyroY axis
%		and with one sensor attached to the participant (HS01)
%		and other sensor attached to the robot (RS01).
%	Code and data to reproduce the figure is available in \cite{srep2021}.
%       }
%    \label{fig:ts}
%\end{figure}
%%%---------------------------------(FIGURE)------------------------------------

>>>>

\section*{Introduction}
The complexity of human movement arises from the process of mastering redundant (bio)mechanical degrees of freedom (DOF) to successfully accomplish any given motor task (Bernstein, 1967).
Such DOF are independent coordinates to uniquely describe the configurations of the system which provides flexibility and stability to adapt to a change environment conditions but that leads many possible combinations (Newell and Vaillancourt, 2001).
Hence, human movement complexity can be seen as a balance between the required DOF "to generate a stable (persistent) and flexible (variable) behavioural output in response to changing intentions and dynamic environmental conditions" \cite{davids2003}.
Consequently, one can see much variability in human movement even in the simplest of movements.
While the analysis of movement of variability is becoming increasingly popular as a diagnostic tool or skill performance evaluation, there remain challenges in terms of defining the most appropriate methods and parameters to apply.
In part, these challenges stem from the fact that the identification movement variability requires analysis of signals which are time-series of $1-$dimension in $\mathbb{R}$ which are noisy, nonlinear and non-stationary \cite{gomezgarcia2014}.
Further problems arise from assigning a plausible locus of control to the movements, e.g., in order to determine whether variability is the result of deliberative control by the performer or whether it arises from exogenous or endogenous disturbances.
For this paper, our focus is on the analysis of signals; specifically, in terms of objectively quantifying variability of lower dimension signals using time-series analysis.

Methods for nonlinear time series analysis generally involve the estimation of the embedding parameters ($m$ embedding dimension and $\tau$ embedding delay) to reconstruct the state space, where an $n$-dimensional is reconstructed state space using $1-$dimensional time series \cite{Quintana-Duque2012, Quintana-Duque2016, sama2013, frank2010, gomezgarcia2014, marwan2011, stergiou2011}.
Key to the selection of these parameters is the need to preserve topological properties of an unknown $M$-dimensional state space \cite{takens1981}.
An common approach to the construction of these state spaces involves Recurrence Plots (RPs), a graphical representation of black and white dots, which shows recurrent patterns of a $n$-dimensional system.
While RPs provide a human interpretable picture of the system, these require further analysis to allow the properties of that system to be quantified, and so Recurrence Quantification Analysis (RQA) can be applied.
However, the estimation of the embedding parameters for RQA remains an open problem (Bradley et al. 2015 \cite{bradley2015}).
There is no agreed method to estimate embedding parameters for RQA and other nonlinear methods \cite{bradley2015} because time series are system-dependent, i.e., these rely on the initial conditions, and on the configuration and behaviour of the system.
This means that, unless one holds all of the influencing variables constant, embedding parameters computed for one instance may not apply to another instance.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Methods}
\subsection*{State Space Reconstruction}
The method of state space reconstruction \cite{packard1980, takens1981} has been applied in many disciplines \cite{aguirre2009, stergiou2011, frank2010, sama2013, Quintana-Duque2016}.
The method of state space reconstruction is based on uniform time-delay embedding methodology which is a simple matrix implementation that can reconstruct an unknown $d-$dimensional manifold $M$ from a scalar time series (e.g. one-dimensional time series in $\mathbb{R}$).
A manifold, in this context, is a multidimensional curved surface within a space (e.g. a saddle) \cite{guastello-gregson2011}.

The use of a scalar time series is the main advantage of the method of state space reconstruction which in essence preserve dynamic invariants such as correlation dimension, fractal dimension, Lyapunov exponents, Kolmogorov-Sinai entropy and detrended fluctuation analysis \cite{bradley2015, Quintana-Duque2012, Quintana-Duque2013, Quintana-Duque2016, krakovska2015}.
However, selecting appropriate embedding parameters which are sued to apply the state space reconstruction is still an open challenge for which we present introductions for the methodologies to compute such embedding parameters.
With that in mind, in the following subsections, we describe in more detail the state space reconstruction theorem (RSSs), uniform time-delay embedding theorem (UTDE), false nearest neighbours (FNN) and average mutual information (AMI).

@@

Following the notation employed in \cite{casdagli1991, garland2016, gibson1992, uzal2011, uzal2010, takens1981}, state space reconstruction is defined by:

@@

where $s$, $s: A \rightarrow M$ given that $A \subseteq \mathbb{R}$ and $M \subseteq \mathbb{R}^d$, represents a trajectory which evolves in an unknown $d-$dimensional manifold $M$, $f: M \rightarrow M$ is an evolution function and $f^t$, with time evolution $t \in \mathbb{N}$, is the $t$-th iteration of $f$ that corresponds to an initial position $s(0) \in M $ \cite{takens1981}.
%$f$ is a smooth dynamical system, where smooth means a continuously differentiable system (e.g. derivatives exist at all points) \cite{guastello-gregson2011}.
Then, a point of a one-dimensional time series $x(t)$ in $\mathbb{R}$, can be obtained with

@@

Reconstructed state space can then be described as an $n-$dimensional state space defined by $y(t)=\Psi[\boldsymbol{X}(t)]$ where $\boldsymbol{X}(t) = \{ x(t), x(t-\tau) , ...,x(t - (m-1)\tau  ) \}$ is the uniform time-delay embedding with a dimension embedding $m$ and delay embedding $\tau$ and $ \Psi: \mathbb{R}^m \rightarrow \mathbb{R}^n$ is a further transformation of dimensionality (e.g. Principal Component Analysis,

@@

With that in mind, uniform time-delay embedding, $\boldsymbol{X}(t)$, defines a map $\Phi: M \rightarrow \mathbb{R}^m$ such that $\boldsymbol{X}(t) = \Phi(s(t))$, where $\Phi$ is a diffeomorphic map \cite{takens1981} whenever $\tau > 0$ and $m > 2d_{box}$ and $d_{box}$ is the box-counting dimension of $M$ \cite{garland2016}.
%"Given two manifolds $M$ and $N$, a differientiable map $f: M \rightarrow N$ is called diffeomorphic if it is one-to-one correspondence and its inverse $f^{-1}: N \rightarrow M$ is differientiable as well \cite{wiki:diffeomorphic}".
Then, if $\Phi$ is an embedding of evolving trajectories in the reconstructed space then a composition of functions represented with $F^t$ is induced on the reconstructed state space determined:

@@

With this in mind, an embedding is defined as "a smooth one-to-one coordinate transformation with a smooth inverse" and the total reconstruction map is defined as $ \Xi = \Psi \circ \Phi $ \cite{casdagli1991}.

@@

		Minimum delay embedding values with AMI's method and minimum dimension embedding values with Cao's method (B), and Recurrence Plots and its patterns (C).

@@

For example, we compute the AMI for two time series: A) the $x$ solution of the Lorenz system, and B) a noise time series using a normal distribution with mean zero and standard deviation equal to one.
From Fig~\ref{fig01}(b4, b8), it can then be concluded that the amount of knowledge for any noise time series is zero for which the first minimum embedding parameter is $\tau_0=1$.
On the contrary, the first minimum of the AMI for the chaotic time series is $\tau_0=17$ which is the value that maximize the independence between $x(n)$ and $x(n+\tau)$ in the reconstructed state space \cite{bradley2015}.
Similarly as Cao's algorithm negatives, AMI's algorithm is not an exception for negatives, which are worthwhile to mention for further investigations.
For instance, (i) is not clear why the choose of the first minimum of the AMI is the minimum delay embedding parameter \cite{kantz2003} and (ii) the probability distribution of the AMI function is computed with the use of histograms which depends on a heuristic choice of number of bins for which AMI depends on partitioning \cite{garcia2005e71}.

@@

%*******************************************************************************
\subsection*{Recurrence Quantifications Analysis (RQA)}
Originally, Zbilut et al. \cite{zbilut1992} proposed metrics to investigate the density of recurrence points in RPs, then histograms of lengths for diagonal lines in RPs were studied by \cite{trulla1996} which were the introduction to the term recurrence quantification analysis (RQA) \cite{marwan2008}.
RQA has been applied in many fields such as life science, engineering, physics, and others \cite{marwan2008}. Particularly in human movement to investigate noise and complexity of postural control \cite{rhea2011}, postural control \cite{apthorp2014} or interpersonal coordination \cite{duran2017}.
The success of RQA is not only due to its simple algorithmic implementation but also to its capacity to detect tiny modulations in frequency or phase which are not detectable using standard methods e.g. spectral or wavelet analysis \cite{marwan2011}, and that RQA's metrics are quantitatively and qualitatively independent of embedding dimension which is verified experimentally by \cite{iwanski1998}.
RQA metrics comprehend percentage of recurrence, percentage of determinism, ratio, Shannon entropy of the frequency distributions of the line lengths, maximal line length and divergence, trend and laminariy \cite{marwan2007, marwan2015}.
For this work, we considered only four RQA metrics, due to its consistency with our preliminary experiments, which are described below.
Such metrics are computed the nonlinearTseries R package \cite{nonlinearTseries2016}.

@@

With that in mind, this work explores the sensitivity and robustness of the window size of time series, embedding parameters for RSS with UTDE and recurrence threshold for RP and RQA in order to gain a better insight into the underlying time series collected from inertial sensors in the context of human-humanoid imitation activities.
%Such changes of recurrence threshold values can modify the patterns in RPs and therefore the values of RQA metrics.
%We therefore computed 3D surfaces to explore the sensibility and robustness of embedding parameters and recurrence threshold in RQA  metrics.
%Following the same methodology of computing 3D surfaces, we also considered variation of window length size to present RQA metrics dependencies with embedding parameters, recurrence thresholds and window length size.
%Zbilut et al. \cite{zbilut1992} established RQA metrics with the aim of determining embedding parameters, their method consisted on creating 3D surfaces with RQA metrics with an increase of embedding parameters ($m$ and $\tau$), then Zbilut et al. \cite{zbilut1992} explored fluctuations and gradual changes in the 3D surfaces that provide information about the embeddings. Much recently, Marwan et al. \cite{marwan2015} created 3D surfaces for visual selection of not only embedding parameters but also recurrence thresholds.
%Following same methodologies, we explored the stability and robustness of RQA metrics (REC, DET, RATIO and ENTR) using 3D surfaces by an unitary increase of the pair embedding parameters ($0 \ge m \le 10$, $0 \ge \tau \le 10$) and a decimal increase of 0.1 for recurrence thresholds ($ 0.2 \ge \epsilon \le 3 $) (Fig.~\ref{fig:topo_rqas}).

@@

\subsection*{Ethics}
All participants provided informed consent forms prior to participation in the experiment. Hence, the experiments were conducted in November 2016 and participants confirmed reading and understanding the participant information sheet of the experiments and were able to withdraw from the experiment at any time without giving any reason.
The design of the experiments adhered to the University of Birmingham regulations, data were anonymised and videos were stored only on a personal computer in accordance with the Data Protection Act 1998.
%Refer to Appendix D for further information about the ethics, online participation information sheets and experiment check list.

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal and vertical arm movements performed by the humanoid robot in the following conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) and faster (HF) speed (Figure~\ref{fig02}a1), and
(ii) ten repetitions of vertical arm movement at normal (VN) and faster (VF) speed (Figure~\ref{fig02}a3).
The normal and faster speed of arm movements is defined by the duration in number of samples of one repetition of NAO's arm movements.
We select NAO's arm movements duration to distinguish between normal and faster arm movements as NAO's movements have less variation for such repetitive movements.
The duration for one repetition of the horizontal arm movement at normal speed, HN, is 5 seconds considering that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds which is around 300 samples of data.
For vertical arm movement at faster speed, VF, each repetition lasts about 2.4 seconds which correspond to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series for axes Z and Y of the gyroscope sensors with four window lengths: 2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) and 15-sec (750-samples).

\subsection*{Data collection from inertial measurement units} 
\label{sec:experiment:subsec:imu}
To give insight to the research questions, we considered various conditions of time series collected for this work:
\begin{itemize}
\item Three levels of smoothness for normalised data (e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg stands for Savitzky-Golay filter with two different filter lengths (29 and 159) and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal} and zmuv is zero mean unit variance.
\item four arm movement activities with two velocities: horizontal normal (HN), horizontal faster (HF), vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}
%After the data collection, raw time series were windowed, normalised and smoothed. 
Due to space limitations and to have simple visualisation, we only present 10-sec (500 samples) window length time series for three participants (p01, p01 and p03) performing horizontal arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig02}C).

\subsubsection*{Raw data}
%Considering the work of \cite{shoaib2016} which provided evidence of an improvement in recognition activities when combining data from accelerometer and gyroscope.
We focus our analysis from data of the accelerometer and gyroscope of the NeMEMsi sensors \cite{Comotti2014} and leave the data of the magnetometer and quaternions for further investigation because of their possible variations with regard to magnetic disturbances \cite{shoaib2016}.
Data from the accelerometer is defined by triaxial time series $A_x(n)$, $A_y(n)$, $A_z(n)$ which forms the matrix $\boldsymbol{A}$ (Eq.~\ref{eq:AG}), and the same for data from the gyroscope which is defined by triaxial time-series of $G_x(n)$, $G_y(n)$, $G_z(n)$ representing the matrix $\boldsymbol{G}$ (Eq.~\ref{eq:AG}).
Both triaxial time series of each sensor, $a$ and $g$, are denoted with its respective axes subscripts $x,y,z$, where $n$ is the sample index and $N$ is the same maximum length of all axes for the time series.
%Matrices  $\boldsymbol{A}$ and $\boldsymbol{G}$ are represented as follows
%%---------------------------------(EQUATION)----------------------------------
\begin{equation}\label{eq:AG}
\boldsymbol{A} =
\begin{pmatrix}
  A_x(n) \\
  A_y(n) \\
  A_z(n)
\end{pmatrix}
=
\begin{pmatrix}
 a_x(1),a_x(2),\dots,a_x(N) \\
 a_y(1),a_y(2),\dots,a_y(N) \\
 a_z(1),a_z(2),\dots,a_z(N) 
\end{pmatrix},
%\end{equation}
%%---------------------------------(EQUATION)-----------------------------------
%and 
%%---------------------------------(EQUATION)-----------------------------------
%\begin{equation}\label{eq:G}
\boldsymbol{G} =
\begin{pmatrix}
 G_x(n) \\
 G_y(n) \\
 G_z(n)
\end{pmatrix}
=
\begin{pmatrix}
 g_x(1),g_x(2),\dots,g_x(N) \\
 g_y(1),g_y(2),\dots,g_y(N) \\
 g_z(1),g_z(2),\dots,g_z(N) 
\end{pmatrix}.
\end{equation}
%%---------------------------------(EQUATION)-----------------------------------
